---
title: "Analysis"
author: "Steven Herrera"
date: "9/15/2020"
fontsize: 12pt
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=1.5cm"
output: 
  pdf_document:
     latex_engine: xelatex
     number_sections: true
---

```{r package_test, include=FALSE}
# Validate that all necessary packaged have been downloaded, install otherwise or throw err package DNE
pkgTest <- function(x) {
  if (!require(x,character.only = TRUE))
    {
    install.packages(x,repos = "http://cran.r-project.org", dep=TRUE)
    if(!require(x,character.only = TRUE)) stop("Package not found")
    }
}
pkgTest("tidyverse")
pkgTest("cowplot")
pkgTest("reticulate")
pkgTest("ggfortify")
pkgTest("ggplot2")
pkgTest("dplyr")
pkgTest("tidyr")
pkgTest("broom")
pkgTest("car")
pkgTest("olsrr")
pkgTest("corrplot")
pkgTest("reshape2")
pkgTest("caret")
pkgTest("gridExtra")
pkgTest("gridGraphics")
pkgTest("grid")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
ggplot2::theme_set(new = theme_bw())
library(tidyverse)
library(cowplot)
library(reticulate)
library(ggfortify)
library(ggplot2)
library(dplyr)
library(tidyr)
library(broom)
library(car)
library(olsrr)
library(corrplot)
library(reshape2)
library(caret)
library(gridExtra)
library(gridGraphics)
library(grid)
#py_install("pandas")
```

```{r load_data, include=FALSE}
raw_stress_data <- read_csv("/Volumes/G-DRIVE mobile/CaseStudy2/amusement_and_stress.csv") %>% 
  mutate(label = as.factor(label))
final <- read_csv("/Volumes/G-DRIVE mobile/CaseStudy2/final.csv") %>% 
  mutate(label = as.factor(label))
```

# Introduction

Wearable technologies have become increasingly prevalent in recent years, with public interest taking note of the accessibility, health interests, and innovativeness [[1]][Bibliography], especially with products such as the Apple Watch. While these devices are commonly marketed as a valuable communication interface for their users, they also provide a wide array of health monitoring features, such as electrocardiography (ECG) [[2]][Bibliography], heart rate variability (HRV) [[3]][Bibliography], and accelerometry (ACC) [[4]][Bibliography]. Health-conscious consumers have increasingly [[1,5,6]][Bibliography] adopted wearables, with [some %] of consumers citing health monitoring as a primary motivating factor for wearable use. 



[citing statistics about how many people wear an Apple watch etc etc] 

Stress is a critical mental and physical health issue affecting all members of the population at some level. It is well-established that stress leads to a number of long-term health conditions, ranging from headaches and sleep issues to an increased risk of cardiovascular disease [[2]][Bibliography]. 

For this reason, researchers have been interested in establishing and interpreting models to predict the affective state of stress using information from wearable devices. [Researcher..] created a [...] model to [...] that achieved [...] results on [...] data. 

A challenge in the detection of stress from wearable data is its physiological similarity to the state of amusement. The two states share [these characteristics], which may be difficult to discern in biological monitoring data. For that reason, this paper will focus on building a model to discern between a state of stress and a state of amusement using the WESAD database. We expand upon previous work by performing more extensive feature engineering, addressing missing data considerations with a K-Nearest Neighbors upsampling strategy, and improving predictive accuracy using a stacking ensemble method.


1. Determine whether sensor data are useful in discriminating between stress and amusement conditions
2. If sensor data are useful in discriminating between stress and amusement, describe how (e.g., is higher blood pressure related to a greater likelihood of stress vs amusement)
3. Determine which types of sensor data are most useful in discriminating between stress and amusement, alone or in combination
4. Determine whether we can detect stress (vs amusement) only using the wrist-worn wearable
5. Quantify the heterogeneity across individuals in the response to stress versus amusement
























# Data

Raw sensor data was recorded on seventeen subjects that participated in the study. However, data of two of the subjects had to be discarded due to sensor malfunction. The data was recorded for each of the subjects with two different devices: RespiBAN, a chest-worn device, and Empatica E4, a wrist-worn device. Within this data analysis, we focused on the synchronised raw sensor data labels, which was created by subjects performing a double tapping gesture with their non-dominant hand on their chest. In summary, the below predictors were included for each of the subjects in the study:

```{r}
# display predictors
```

## Missing Data


## Feature Engineering

The raw signal data was subsetted into 100-second rolling windows in Python, referenced in the Appendix (what figure). It was necessary to subset the data into the 100-second windows in order to calculate the features for each of the windows, which is a methodology practiced by (AUTHORS) in the (WESAD PAPER). Additionally, the method of subsetting signal data into windows is performed by Apple when the Apple Watch measures the mean ECG of its users. 100-second windows were chosen because of the necessity to extract features that made sense, as with not enough data, the features extracted would not make sense in the scope of the problem. (i.e. add more about how it wouldn’t make sense intuitively )

#### should we connect what was said about the window stuff (above) with this paragraph?

Feature Extraction was performed on the raw signal data, in order to derive important aspects concerning the data that are often used by medical professionals to evaluate the health of their patients (cite this). Feature extraction was performed for some of the raw sensor data with the Python package neurokit2 (CITE the package), and by Python functions provided in the Appendix with (CITE THIS). 

## Exploratory Data Analysis



















### Correlations

```{r correlations, warning=FALSE}
# remove columns w NAs, remove subject, remove identifier column X1
final_na <- final[, colSums(is.na(final)) != nrow(final)] %>% dplyr::select(-subject, -X1)

# remove label (the amusemet or stress identifier variable) to do correlation analysis
final_na$label <- NULL
# rename
d <- final_na
# apply correlations
zv <- apply(final_na, 2, function(x) length(unique(x)) == 1)
dfr <- final_na[, !zv]
n=length(colnames(dfr))
corr.d <- cor(dfr[,1:n],use="complete.obs")
# NAs on the diagonal
corr.d[ lower.tri( corr.d, diag = TRUE ) ] <- NA

# # sort pairs of correlated variables
# m <- melt(corr.d)
# m <- m[order(- abs(m$value)), ]
# # get rid of NAs
# m <- na.omit(m)
# # view the highest correlated: correlation vals > 0.75
# (check <- as.data.frame(m[ which( m$value > 0.80 ), ]))

# plot the correlations (it looks messy)
corrplot(corr.d, type = "upper", diag = FALSE , method = "color",
         tl.pos = "td", tl.cex = 0.25)
```

```{r}
# recall corr without the diagonal transformation
corr.d <- cor(dfr[,1:n],use="complete.obs")
# determine the variables most important with cutoff at 0.7
highlyCorrelated <- findCorrelation(corr.d, cutoff=(0.7),verbose = FALSE)
important_var <- colnames(corr.d[,-highlyCorrelated])
non_important_var <- colnames(corr.d[,highlyCorrelated])

# recreate the correlation plot
final_select <- final_na %>%
  select(c(important_var))

# apply correlations
zv1 <- apply(final_select, 2, function(x) length(unique(x)) == 1)
dfr1 <- final_select[, !zv1]
n=length(colnames(dfr1))
corr.d1 <- cor(dfr1[,1:n],use="complete.obs")
# NAs on the diagonal
corr.d1[ lower.tri( corr.d1, diag = TRUE ) ] <- NA

corrplot(corr.d1, type = "upper", diag = FALSE , method = "color",
         tl.pos = "td", tl.cex = 0.25,
         title = "Variables with low correlation (rho 0.7)")
```

```{r}
# recreate the correlation plot
final_select1 <- final_na %>%
  select(c(non_important_var))

# apply correlations
zv1 <- apply(final_select1, 2, function(x) length(unique(x)) == 1)
dfr1 <- final_select1[, !zv1]
n=length(colnames(dfr1))
corr.d12 <- cor(dfr1[,1:n],use="complete.obs")
# NAs on the diagonal
corr.d12[ lower.tri( corr.d12, diag = TRUE ) ] <- NA

corrplot(corr.d12, type = "upper", diag = FALSE , method = "color",
         tl.pos = "td", tl.cex = 0.25)
```





































# Methodology

# Results

# Cross-Validation // Sensitivity Analysis(?)

# Conclusion & Limitations

# Appendix


## Bibliography

1. Lee, Sang Yup, and Keeheon Lee. “Factors That Influence an Individual's Intention to Adopt a Wearable Healthcare Device: The Case of a Wearable Fitness Tracker.” Technological Forecasting &amp; Social Change, vol. 129, 2018, pp. 154–163., doi:10.1016/j.techfore.2018.01.002.

2. Murphy, Jane Louise, et al. “The Use of Wearable Technology to Measure Energy Expenditure, Physical Activity, and Sleep Patterns in Dementia.” Alzheimer's &amp; Dementia: The Journal of the Alzheimer's Association, vol. 11, no. 7, 2015, pp. P188–P188., doi:10.1016/j.jalz.2015.07.165.

3. Nixon, Jim, and Rebecca Charles. “Understanding the Human Performance Envelope Using Electrophysiological Measures from Wearable Technology.” Cognition, Technology &amp; Work, vol. 19, no. 4, 2017, pp. 655–666., doi:10.1007/s10111-017-0431-5.

4. Prieto, L. P., et al. “Multimodal Teaching Analytics: Automated Extraction of Orchestration Graphs from Wearable Sensor Data.” Journal of Computer Assisted Learning, vol. 34, no. 2, 2018, pp. 193–203., doi:10.1111/jcal.12232. Accessed 29 Sept. 2020. 

5. Jakicic, John M., et al. “Effect of Wearable Technology Combined With a Lifestyle Intervention on Long-Term Weight Loss: The IDEA Randomized Clinical Trial.” JAMA : the Journal of the American Medical Association, vol. 316, no. 11, 2016, pp. 1161–1171., doi:10.1001/jama.2016.12858.

6. Bai, Yang, et al. “Comparative Evaluation of Heart Rate-Based Monitors: Apple Watch vs Fitbit Charge HR.” Journal of Sports Sciences, vol. 36, no. 15, 2017, pp. 1734–1741., doi:10.1080/02640414.2017.1412235.




5. Schmidt, Philip, et al. “Introducing WESAD, a Multimodal Dataset for Wearable Stress and Affect Detection.” Proceedings of the 2018 on International Conference on Multimodal Interaction - ICMI '18, 2018, doi:10.1145/3242969.3242985. 

3. 

