---
title: "Analysis"
author: "Steven Herrera"
date: "9/15/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Context: N=15 participants with a mean age 27.5 +- 2.4 years; 12 were male and 3 were were female. 
Goals: Differentiating between states: neutral, stress, and amusement.

#### RespiBAN

RespiBAN: equipped with (700Hz) sensors to measure ACC and RESP; has four "analog ports": ECG, EDA, EMG, and TEMP.

- ECG data was recorded via a standard 3-point ECG.

#### Empatica E4

Empatica E4: worn on non-dominant hand; recorded BVP (64Hz), EDA (4Hz), TEMP (4Hz), & ACC (32Hz).





```{r cars}
pickle_sample <- read_csv("/Users/stevenherrera/Downloads/CaseStudy2/data/un_pickled_data/sampled_pickle.csv")
```

```{r}
# X1: Sample ID
# ACC, ECG, EDA, EMG, RESP, TEMP are all RespiBAN CHEST data info
# Label:  ID of the respective study protocol condition, sampled at 700 Hz
# ^  0 = not defined / transient, 1 = baseline, 2 = stress, 3 = amusement, 4 = meditation
# ^  5/6/7 = should be ignored in this dataset

pickle_sample %>%
  count(Label)

```

